---
title: "R Markdown Presentation & Plotly Week 3 Project"
author: "Alexander M Fisher"
date: "30/12/2020"
output: ioslides_presentation
---


## Intro to Project:

Create a web page presentation using R Markdown that features a plot created with Plotly. Host your webpage on either GitHub Pages, RPubs, or NeoCities. Your webpage must contain the date that you created the document, and it must contain a plot created with Plotly. We would love to see you show off your creativity! 

Review Criteria - The rubric contains the following two questions:

- Does the web page feature a date and is this date less than two months before the date that you're grading this assignment?
- Is the web page a presentation and does it feature an interactive plot that appears to have been created with Plotly?


## Getting and Preparing Data:

In this project to demonstrate the use of plotly I will redo an exploratory box plot done 
in a previous assignment using plotly. That is the regression model project in which we explored the relationship between mpg 
and transmission type from the `mtcars` data set. 

```{r}
suppressMessages(library(plotly))
data("mtcars")
data <- mtcars
data$am <- factor(data$am, labels = c("auto","manual"))
data$cyl <- factor(data$cyl)
print(str(data))
```

## Plotly Boxplot:

```{r fig.dim=c(6,4)}
box_plot <- plot_ly(data = data, x=data$am, y=data$mpg, color = data$am, 
                    type = "box")
options(warn = -1)
box_plot
```

